### 2020-06-23 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
explored sampsize fucntion

mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High"))) %>% 
  # mutate(dnbr_qualitative = case_when(
  #   dnbr < 0.1 ~ "unburned",
  #   dnbr < 0.27 ~ "low",
  #   dnbr < 0.66 ~ "moderate",
  #   dnbr > 0.66 ~ "high")) %>% 
  # mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
  #        dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("high", "moderate", "low", "unburned"))) %>% 
  mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
  mutate(three_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% "Moderate" ~ "Moderate",
    mortality_qual %in% c("Low", "None") ~ "Low")) %>% 
  mutate(three_class_mortality = as.factor(three_class_mortality)) %>% 
  mutate(three_class_mortality= fct_relevel(three_class_mortality, 
                                     levels=c("Low", "Moderate", "High")))

@@

@@

#random forest code---------------------------------------------------
for (i in 1:50)
{

@@

  #random forest model
  rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ridge + drainage_association + ndvi, data = train, ntree=1500, importance=T, nscale=0)

@@

  # lists to save model outputs
  test_accuracy_list <- append(test_accuracy_list, test_accuracy)
  write.table(model_importance, file="rfm model tc importance.csv", sep=",",col.names=NA, append=T) 
  write.table(tabb, file = "mcf tc test accuracy table.csv", sep=",",col.names=NA, append=T)
}

>>>>

mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High"))) %>% 
  # mutate(dnbr_qualitative = case_when(
  #   dnbr < 0.1 ~ "unburned",
  #   dnbr < 0.27 ~ "low",
  #   dnbr < 0.66 ~ "moderate",
  #   dnbr > 0.66 ~ "high")) %>% 
  # mutate(dnbr_qualitative = as.factor(dnbr_qualitative),
  #        dnbr_qualitative = fct_relevel(dnbr_qualitative, levels=c("high", "moderate", "low", "unburned"))) %>% 
  mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
  mutate(three_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% "Moderate" ~ "Moderate",
    mortality_qual %in% c("Low", "None") ~ "Low")) %>% 
  mutate(three_class_mortality = as.factor(three_class_mortality)) %>% 
  mutate(three_class_mortality= fct_relevel(three_class_mortality, 
                                     levels=c("Low", "Moderate", "High"))) %>% 
  mutate(tc_mortality = as.factor(tc_mortality))

@@

### Random Forest forced sample sizes

https://stats.stackexchange.com/questions/457451/how-r-randomforest-sampsize-works

tc_mortality 1 = type converted
tc_mortality 0 = "not_tc"

using for loop to generate several graphs so its easier to compare them
```{r}

for (i in 1:10) {

rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + drainage_association + ndvi, 
                    data = mcf, 
                    sampsize = c('1' = 99, '0' = 99), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
# OOB and importance
rfm
importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

```
variable importance seems stable AND class accuracy is high and equal between classes!!

#### ROC and AUC
```{r}
# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
rfm_roc <-  roc(mcf$tc_mortality, rfm$votes[,1])
plot(rfm_roc) # plot comes out a little weird
auc(rfm_roc)

## PDP v. ALE

```{r}
#PDP
a <- partialPlot(rfm, as.data.frame(mcf), ndvi, "0") #0=not type converted
b <- partialPlot(rfm, as.data.frame(mcf), ndvi, "1", col="red") #1=type converted
lines(a, col="green")

a <- partialPlot(rfm, as.data.frame(mcf), slope, "0") #0=not type converted
b <- partialPlot(rfm, as.data.frame(mcf), slope, "1", col="red") #1=type converted
lines(a, col="green")

a <- partialPlot(rfm, as.data.frame(mcf), sr, "0") #0=not type converted
b <- partialPlot(rfm, as.data.frame(mcf), sr, "1", col="red") #1=type converted
lines(a, col="green")

#ale (see detailed notes in ALE Plots code chunk below; not working right now)
train_x <- mcf[-which(names(mcf) == "tc_mortality")]
ndvi_predictor_notTC <- Predictor$new(rfm, data = train_x, y = mcf$tc_mortality, type = "prob", class = "0")
ndvi_not <- plot(FeatureEffect$new(ndvi_predictor_notTC, feature = "ndvi")) + ggtitle("not TC")

#predict values for the other mortality classes for the same variable
ndvi_predictor_tc <- Predictor$new(rfm, data = train_x, type = "prob", class = "1")
sr_moderate <- plot(FeatureEffect$new(ndvi_predictor_tc, feature = "ndvi")) + ggtitle("TC")
```
so looking at the 3 most important variables for type conversino plots, all of the probabilities are below zero. only the not type converted stands have probabilities above 0, with some values of the explanatory variable having very high probabilities. So what this tells me is the model CANNOT distinguish type converted stands from non type converted stands.

### Random Forest and forced sample sizes with training and test data

hmmmm. not sure how to do this with training and test data since the train/test data frames will have different quantities of the type conversion classes each time. if i do it the way I have been doing it below, then variable importance is too unstable. even using my equal distribution sample size code and adding in the sampsize function results in unstable variable importance. 
```{r}
#train and test data 
sample <- sample(1:nrow(mcf), size = 0.75*nrow(mcf))
train <- mcf[sample, ]
test  <- mcf[-sample, ]

for (i in 1:10) {
rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + vpd_jan + drainage_association + ndvi, 
                    data = mcf, 
                    sampsize = c('1' = 99, '0' = 99), 
                    ntree=1500, 
                    importance=T,
                    replace = T)
rfm

#plot to visualize variable importance
varImpPlot(rfm)
#model_importance <- importance(rfm)
importance(rfm)
}
#predcit test data
pp<-predict(rfm, test)

#test data accuracy
tabb<- table(pp, test$tc_mortality)
acc<-1-sum(diag(tabb))/sum(tabb) # % missclassified 
test_accuracy <-(1-acc)*100 #model accuracy
```

@@

#random forest code---------------------------------------------------
for (i in 1:15)
{

@@

  #random forest model
  rfm <- randomForest(tc_mortality ~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ridge + drainage_association + ndvi, data = train, ntree=1500, importance=T, nscale=0, sampsize = c('tc' = 75, 'not_tc' = 75), replace = T)

@@

  # lists to save model outputs
  # test_accuracy_list <- append(test_accuracy_list, test_accuracy)
  # write.table(model_importance, file="rfm model tc importance.csv", sep=",",col.names=NA, append=T) 
  # write.table(tabb, file = "mcf tc test accuracy table.csv", sep=",",col.names=NA, append=T)
}
